% !TeX encoding = UTF-8
% !TeX spellcheck = en_US
% !TeX root = slackforceDocumentation.tex

\chapter{Introduction}

For slackliners it is often important to know the physical forces involved in slacklining, for example to estimate the suitability of potential slackline anchors. For static conditions, some well known calculations exist, to get the force on a slackline anchor from the sag of the slackline and the weight of the slackliner. However, it is often not easy to estimate the sag of a slackline precisely. Therefore it would be nice to have other measurement methods as well. Even more complicated is the estimation of the peak force in a slackline during dynamic movements like jumping.

The idea of the app ``Slackforce'' is to provide an easy to use solution for measuring and calculating forces in a slackline. This document describes the physical background and derivation of formulas and calculations used within this app. The app contains the following three main functionalities:

\begin{description}
	\item[Measure Force:] Measure the pretension of a slackline from the sound you hear when slapping the line with your finger.
	\item[Static Calculations:] Calculate the forces on a slackline at static conditions resulting from the geometric setup of the line.
	\item[Dynamic Simulation:] Simulate the movement and forces of a slackline during dynamic conditions like bouncing or jumping.
\end{description}

This document is divided into three chapters corresponding to the app functionalities. Note that the presented calculations are experimental and might contain errors. Use the formulas careful and do not fully trust the calculated values. The creator of this document is not responsible for any harm resulting from the use of the presented calculations and formulas.